---
title: "MOCHI Kids — new kid data (attention-filtered)"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M %Z')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: hide
    self_contained: true
params:
  refresh: false          # set TRUE to re-pull from Mongo via fetch_data.py
  focus_date: null        # "YYYY-MM-DD"; null = most recent day in the data
  spam_share: 0.95        # sessions choosing one position >= this share are pure mash-tests
  fast_ms: 400            # RT below this = a tap too fast to be a real decision
  mash_run: 8             # a run of this many identical positions = mashing
  chance: 0.3333          # 3-image oddity
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 9, fig.height = 4.6, fig.align = "center",
  dev = "png", dpi = 150,
  fig.path = here::here("analysis", "figures", "kid_")
)
library(tidyverse)
library(scales)
library(here)

# Brand-neutral, CVD-safe palette (from the dataviz reference palette)
PAL <- c(blue = "#2a78d6", orange = "#eb6834", green = "#008300",
         violet = "#4a3aa7", grey = "#8a8a86")
GOOD <- "#0ca30c"; BAD <- "#d03b3b"
TIER_LEVELS <- c("training", "warmup", "familiar", "novel", "catch")

theme_set(theme_minimal(base_size = 13) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"),
        strip.text = element_text(face = "bold")))
```

```{r load}
if (isTRUE(params$refresh)) system2("python3", c(here("analysis", "fetch_data.py")))

sessions <- read_csv(here("analysis", "data", "sessions.csv"), show_col_types = FALSE)
trials   <- read_csv(here("analysis", "data", "trials.csv"),   show_col_types = FALSE)

# Which day are we analyzing? Default: the most recent day with data.
sessions <- sessions %>% mutate(day = as.Date(finishedAt))
focus_day <- if (is.null(params$focus_date)) max(sessions$day, na.rm = TRUE) else as.Date(params$focus_date)

focus_sessions <- sessions %>% filter(day == focus_day)
focus_ids <- focus_sessions$participantID
```

Analyzing **`r length(focus_ids)`** session(s) from **`r as.character(focus_day)`**.

```{r flag-engagement}
# Per-trial attention filter. A trial is "valid" (a real decision) unless it is
# either too fast to be a choice, or part of a run of identical positions
# (button-mashing the same spot).
ft <- trials %>%
  filter(participantID %in% focus_ids) %>%
  arrange(participantID, trial_index) %>%
  group_by(participantID) %>%
  mutate(run_id = cumsum(replace_na(chosen_display_pos != lag(chosen_display_pos), TRUE))) %>%
  group_by(participantID, run_id) %>%
  mutate(run_len = n()) %>%
  ungroup() %>%
  mutate(
    fast  = rt < params$fast_ms,
    mash  = run_len >= params$mash_run,
    valid = !fast & !mash,
    tier  = factor(tier, levels = TIER_LEVELS),
    outcome = if_else(correct, "correct", "incorrect")
  ) %>%
  left_join(select(focus_sessions, participantID, consent_age), by = "participantID")

# Session-level roll-up + a data-quality verdict.
sess_qa <- focus_sessions %>%
  mutate(
    pure_mash = max_position_share >= params$spam_share,
    label = sprintf("%s (age %s)", participantID, consent_age),
    verdict = case_when(
      pure_mash                     ~ "EXCLUDE — position spam",
      overall_acc < params$chance & max_run_length > 20 ~ "PARTIAL — disengaged",
      TRUE                          ~ "USABLE"
    )
  )

# Nice ordered participant labels for plotting (drop the pure-mash test from kid plots)
kid_ids  <- sess_qa %>% filter(!pure_mash) %>% pull(participantID)
lab_of   <- setNames(sess_qa$label, sess_qa$participantID)
ft <- ft %>% mutate(who = fct_reorder(lab_of[participantID], as.integer(factor(participantID))))

# Stable color per participant, shared across every plot (color follows the entity).
who_levels <- levels(ft$who)
who_colors <- setNames(unname(PAL[c("blue","orange","violet","green")])[seq_along(who_levels)], who_levels)
```

## Session overview & data-quality verdict

```{r overview}
sess_qa %>%
  transmute(
    session = participantID, age = consent_age,
    n_trials, raw_acc = percent(overall_acc, 1),
    mean_rt_ms = round(mean_rt),
    pos_bias = percent(max_position_share, 1),
    longest_run = max_run_length,
    n_fast = n_fast_lt_500ms,
    verdict
  ) %>%
  knitr::kable(caption = sprintf("Sessions on %s", focus_day))
```

The pure-mash test session is excluded from the kid analyses below; it is a
button-mash QA run (one position, near-instant taps), not a participant.

## Engagement over the session (RT per trial)

The clearest read on a young child: as long as they are deciding, RTs stay up
(seconds). When they check out and tap through, RTs collapse below the
`r params$fast_ms` ms line and the same position repeats. Point shape/color =
whether the answer was correct.

```{r rt-per-trial, fig.height=4.8}
ggplot(filter(ft, participantID %in% kid_ids),
       aes(trial_index, rt)) +
  geom_hline(yintercept = params$fast_ms, linetype = "dashed", color = PAL["grey"]) +
  geom_line(alpha = 0.25, color = PAL["grey"]) +
  geom_point(aes(color = outcome, shape = outcome), size = 2) +
  scale_y_log10(labels = comma) +
  scale_color_manual(values = c(correct = GOOD, incorrect = BAD)) +
  scale_shape_manual(values = c(correct = 16, incorrect = 4)) +
  facet_wrap(~ who, ncol = 1) +
  labs(title = "Response time across the session",
       subtitle = sprintf("Dashed line = %d ms 'too fast to be a choice' threshold", params$fast_ms),
       x = "Trial number", y = "RT (ms, log scale)", color = NULL, shape = NULL)
```

## Rolling accuracy (window = 10 trials)

```{r rolling-acc}
roll <- function(x, k = 10) {
  sapply(seq_along(x), function(i) {
    lo <- max(1, i - k + 1); mean(x[lo:i])
  })
}
ft %>%
  filter(participantID %in% kid_ids) %>%
  group_by(who) %>%
  arrange(trial_index) %>%
  mutate(racc = roll(as.numeric(correct))) %>%
  ungroup() %>%
  ggplot(aes(trial_index, racc, color = who)) +
  geom_hline(yintercept = params$chance, linetype = "dotted") +
  annotate("text", x = 3, y = params$chance + 0.03, label = "chance", size = 3, color = PAL["grey"]) +
  geom_line(linewidth = 1) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_color_manual(values = who_colors) +
  labs(title = "Rolling accuracy", x = "Trial number", y = "Accuracy (10-trial window)",
       color = NULL) +
  theme(legend.position = "top")
```

## Accuracy by tier — raw vs. attention-filtered

Filtering to valid trials strips out mashed responses, so it reflects what the
child did *while actually playing*. `n` above each bar is the number of valid
trials remaining in that tier.

```{r tier-acc, fig.height=5}
raw_tier <- ft %>% filter(participantID %in% kid_ids) %>%
  group_by(who, tier) %>% summarise(acc = mean(correct), n = n(), .groups = "drop") %>%
  mutate(set = "raw")
val_tier <- ft %>% filter(participantID %in% kid_ids, valid) %>%
  group_by(who, tier) %>% summarise(acc = mean(correct), n = n(), .groups = "drop") %>%
  mutate(set = "attention-filtered")

bind_rows(raw_tier, val_tier) %>%
  mutate(set = factor(set, levels = c("raw", "attention-filtered"))) %>%
  ggplot(aes(tier, acc, fill = set)) +
  geom_hline(yintercept = params$chance, linetype = "dotted") +
  geom_col(position = position_dodge(0.8), width = 0.7) +
  geom_text(aes(label = n), position = position_dodge(0.8), vjust = -0.4, size = 3) +
  scale_y_continuous(labels = percent, limits = c(0, 1.05)) +
  scale_fill_manual(values = c(raw = PAL[["grey"]], "attention-filtered" = PAL[["blue"]])) +
  facet_wrap(~ who) +
  labs(title = "Accuracy by tier", subtitle = "dotted = chance; number = valid n",
       x = NULL, y = "Accuracy", fill = NULL) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 30, hjust = 1))
```

## Position choices (mashing check)

A genuinely engaged child spreads taps across all three positions; a
disengaged one collapses onto one.

```{r pos-dist}
ft %>%
  mutate(pos = factor(chosen_display_pos)) %>%
  count(who, pos) %>%
  group_by(who) %>% mutate(share = n / sum(n)) %>% ungroup() %>%
  ggplot(aes(pos, share, fill = who)) +
  geom_col(width = 0.7) +
  geom_hline(yintercept = 1/3, linetype = "dotted") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = who_colors, guide = "none") +
  facet_wrap(~ who) +
  labs(title = "Chosen screen position", subtitle = "dotted = even 1/3 split",
       x = "Position (0 = left, 1 = middle, 2 = right)", y = "Share of trials")
```

## Takeaways

```{r takeaways, echo=FALSE, results='asis'}
n_kid <- length(kid_ids)
n_excl <- sum(sess_qa$pure_mash)
cat(sprintf("- **%d usable kid session(s)**, **%d excluded** as position-spam QA test(s).\n", n_kid, n_excl))
cat("- Engagement, not ability, drives the raw scores — watch the RT-per-trial and position plots, not just overall accuracy.\n")
cat("- The attention filter salvages the engaged portion of a session instead of discarding the whole child.\n")
cat(sprintf("- Chance is %s; catch trials should sit near ceiling — if engaged kids miss them, inspect the catch stimuli.\n", percent(params$chance, 1)))
```
