---
title: "MOCHI Kids — pilot sanity check"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  refresh: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                     fig.width = 8, fig.height = 4.5)
library(tidyverse)
library(scales)
library(here)

if (params$refresh) {
  system2("python3", c(here("analysis", "fetch_data.py")))
}

sessions <- read_csv(here("analysis", "data", "sessions.csv"),
                     show_col_types = FALSE)
trials   <- read_csv(here("analysis", "data", "trials.csv"),
                     show_col_types = FALSE)

# Treat consent_age as factor with sensible order (kids first, adult last).
age_order <- c("2","3","4","5","6","7","8","9","10+","adult")
sessions <- sessions %>% mutate(consent_age = factor(consent_age, levels = age_order))
trials   <- trials   %>% mutate(consent_age = factor(consent_age, levels = age_order),
                                tier = factor(tier, levels = c("training","warmup","familiar","novel")))

n_sessions <- nrow(sessions)
n_trials   <- nrow(trials)
```

> Set `params: refresh: yes` (or knit with refresh=TRUE) to re-pull from Mongo.

`r if (n_sessions == 0) "**No real sessions in the database yet.** Run a few playthroughs (or wait until pilot data lands) and re-knit.\n"`

```{r empty-guard, eval = (n_sessions == 0), echo = FALSE, results = "asis"}
knitr::knit_exit()
```

# Sample

`r n_sessions` sessions, `r n_trials` trials.

```{r}
sessions %>%
  count(consent_age, .drop = FALSE) %>%
  ggplot(aes(consent_age, n)) +
  geom_col(fill = "#6ec1e4") +
  labs(x = "self-reported age", y = "sessions",
       title = "Who's played so far") +
  theme_minimal()
```

# Completion

The trial set has 74 trials. A complete session has `n_trials == 74`.

```{r}
sessions %>%
  ggplot(aes(n_trials)) +
  geom_histogram(binwidth = 1, fill = "#ff6f61") +
  geom_vline(xintercept = 74, linetype = "dashed") +
  labs(x = "trials completed", y = "sessions",
       title = "Completion histogram (dashed = full set)") +
  theme_minimal()

incomplete <- sessions %>% filter(n_trials < 74)
if (nrow(incomplete) > 0) {
  cat("Sessions that didn't finish:\n")
  knitr::kable(incomplete %>% select(participantID, consent_age, n_trials, n_correct))
}
```

# Accuracy by tier

Adult baselines from the README: training≈1, warmup≈1, familiar≥0.95, novel≥0.95.
Chance for 3-AFC is 1/3.

```{r}
acc_by_tier <- trials %>%
  group_by(tier) %>%
  summarise(n = n(), acc = mean(correct, na.rm = TRUE), .groups = "drop")

knitr::kable(acc_by_tier, digits = 3)

trials %>%
  group_by(participantID, consent_age, tier) %>%
  summarise(acc = mean(correct, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(tier, acc, color = consent_age)) +
  geom_hline(yintercept = 1/3, linetype = "dashed", color = "grey50") +
  geom_jitter(width = 0.15, height = 0, size = 3, alpha = 0.8) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(x = NULL, y = "accuracy",
       title = "Per-participant accuracy by tier",
       caption = "dashed line = chance (1/3)") +
  theme_minimal()
```

# Reaction time

Sanity: kids should mostly land between 500ms and ~10s.
Floor (button-mash) and ceiling (stuck) flag UI issues.

```{r}
trials %>%
  filter(!is.na(rt)) %>%
  ggplot(aes(tier, rt / 1000, fill = tier)) +
  geom_violin(alpha = 0.5) +
  geom_jitter(width = 0.1, height = 0, size = 0.5, alpha = 0.4) +
  scale_y_continuous(trans = "log10",
                     breaks = c(0.1, 0.25, 0.5, 1, 2, 5, 10, 30),
                     labels = c("0.1","0.25","0.5","1","2","5","10","30")) +
  labs(x = NULL, y = "RT (seconds, log)",
       title = "RT distribution by tier") +
  theme_minimal() +
  theme(legend.position = "none")

trials %>%
  group_by(tier) %>%
  summarise(median_rt_s = median(rt, na.rm = TRUE) / 1000,
            p25_rt_s    = quantile(rt, 0.25, na.rm = TRUE) / 1000,
            p75_rt_s    = quantile(rt, 0.75, na.rm = TRUE) / 1000,
            n_under_500ms = sum(rt < 500, na.rm = TRUE),
            n_over_15s    = sum(rt > 15000, na.rm = TRUE),
            .groups = "drop") %>%
  knitr::kable(digits = 2)
```

# Per-trial accuracy

If everyone (or no one) gets a specific trial right, the stimulus may be broken
or trivially easy. Compare kid accuracy to the adult baseline.

```{r fig.height = 6}
per_trial <- trials %>%
  group_by(trial_id, tier, dataset, human_avg_adult) %>%
  summarise(n = n(), kid_acc = mean(correct, na.rm = TRUE), .groups = "drop")

per_trial %>%
  ggplot(aes(human_avg_adult, kid_acc, color = tier)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_hline(yintercept = 1/3, linetype = "dotted", color = "grey60") +
  geom_point(size = 2.5, alpha = 0.8) +
  scale_x_continuous(labels = percent_format(), limits = c(0, 1)) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(x = "adult accuracy (from manifest)",
       y = "kid accuracy (this dataset)",
       title = "Per-trial: kids vs adults",
       caption = "dashed = identity, dotted = chance") +
  theme_minimal()

odd <- per_trial %>%
  filter((n >= 3) & (kid_acc <= 1/3 - 0.05 | kid_acc >= 0.99)) %>%
  arrange(kid_acc)
if (nrow(odd) > 0) {
  cat("Trials worth eyeballing (kids near chance OR everyone correct, n>=3):\n")
  knitr::kable(odd, digits = 3)
}
```

# Position bias

If kids systematically tap left/middle/right regardless of where the oddity is,
something's off (visual layout, audio cuing, etc.).

```{r}
trials %>%
  count(chosen_display_pos) %>%
  mutate(prop = n / sum(n)) %>%
  knitr::kable(digits = 3)

trials %>%
  count(chosen_display_pos, correct) %>%
  ggplot(aes(factor(chosen_display_pos), n, fill = correct)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "display position chosen (0=left, 1=mid, 2=right)",
       y = "share", title = "Accuracy by chosen position") +
  theme_minimal()
```

# Schema sanity

Any rows with missing fields that should be there.

```{r}
session_missing <- sessions %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "field", values_to = "n_missing") %>%
  filter(n_missing > 0)

trial_missing <- trials %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "field", values_to = "n_missing") %>%
  filter(n_missing > 0)

cat("session-level missing:\n"); knitr::kable(session_missing)
cat("\ntrial-level missing:\n");  knitr::kable(trial_missing)
```
